Modern Graph Generators: A Survey

### 2018-09-24 generators_community_detection.tex
Sherif Changes

The first attempts to compare community detection algorithms using synthetic graphs proposed the use of random graphs composed by several Erd\"{o}s-R\'{e}nyi subgraphs, connected more internally than externally~\cite{danon2005comparing}. Each of these subgraphs has the same size and the same internal/external density of edges. However, such graphs miss the realism observed in real graphs, where communities are of different sizes and densities. Thus, Lancichinetti, Fortunato and Radicchi (hence LFR)~\cite{PhysRevE.78.046110} propose a class of benchmark graphs for community detection where communities are of different sizes and densities. The generator generates communities of different sizes following a power-law distribution whose parameters can be configured. The degree of the nodes is also sampled from a power-law distribution. Additionally, the generator introduces the concept of the ``mixing factor'', which specifies the fraction of its links connecting nodes belonging to different communities. Such parameter allows to tune the degree of modularity of the generated graph, thus testing the robustness of the algorithms under different conditions. The generation process starts with an empty graph and incrementally fills in the adjacency matrix by obeying the described constraints. Lancichinetti, Fortunato and Radicchi~\cite{PhysRevE.80.016118} extended LFR to support the notion of directed graphs and overlapping communities. Overlapping communities extend the notion of communities by allowing the sharing of vertices, thus a vertex can belong to more than one community. This extended generator allows controlling the same parameters of the previous version, but also the amount of overlap of the generated communities.

Besides synthetic graph generators, Yang and Leskovec~\cite{yang2015defining} proposed the use of real-world graphs with explicit group annotations (e.g., forums in a social network, categories of products, etc.) to infer what they call ``meta-communities'', and use them to evaluate overlapping community detection algorithms. However, a recent study from Hric, Darst and Fortunato~\cite{hric2014community} reveal a loose correspondence between  communities (the authors refer to them as structural communities) and meta-communities.  This result reveals that  algorithms working for structural communities do not work well for finding meta-communities and viceversa, suggesting significantly different underlying characteristics between the two types of communities, which are yet to be identified.

In this regard and to the best of our knowledge, there does not exist a set of generators to generate graphs with meta-communities for community detection algorithm benchmarking. The closest one is the LDBC-SNB data generator~\cite{Erling:2015:LSN:2723372.2742786} thanks to the generation of groups of users in the social network. Even though the generation process does not specifically enforce the generation of groups (meta-communities) for benchmarking community detection algorithms, the study from Prat-Pe\'rez and Dominguez-Sal reveals that these groups are more similar to the real meta-communities than those structural communities generated by the LFR benchmark~\cite{Prat-Perez:2014:CSS:2621934.2621942}.

The differences observed between structural and meta-communities reveal the need of more accurate community definitions tight more specifically to the domain or use case. Current community detection algorithms and graph generators for community detection are stuck to the traditional (and vague) definition of community, assuming that there exists a single algorithm that would fit all the use cases. Thus, future work requires the study of domain-specific community characteristics that can be used to generate graphs with a community structure that accurately resemble that of specific use cases, and thus revealing which are the best algorithms for each particular scenario.

>>>>

The first attempts to compare community detection algorithms using synthetic graphs proposed the use of random graphs composed by several Erd\"{o}s-R\'{e}nyi subgraphs, connected more internally than externally~\cite{danon2005comparing}. Each of these subgraphs has the same size and the same internal/external density of edges. However, such graphs miss the realism observed in real graphs, where communities are of different sizes and densities. Thus, Lancichinetti, Fortunato and Radicchi (hence LFR)~\cite{PhysRevE.78.046110} propose a class of benchmark graphs for community detection where communities are of different sizes and densities. The generator generates communities of different sizes following a power-law distribution whose parameters can be configured. The degree of the nodes is also sampled from a power-law distribution. Additionally, the generator introduces the concept of the ``mixing factor'', which specifies the fraction of its links connecting nodes belonging to different communities. Such parameter allows  the degree of modularity of the generated graph  to be tuned, thus testing the robustness of the algorithms under different conditions. The generation process starts with an empty graph and incrementally fills in the adjacency matrix by obeying the described constraints. Lancichinetti, Fortunato and Radicchi~\cite{PhysRevE.80.016118} extended LFR to support the notion of directed graphs and overlapping communities. Overlapping communities extend the notion of communities by allowing the sharing of vertices, thus a vertex can belong to more than one community. This extended generator allows controlling the same parameters of the previous version, but also the amount of overlap of the generated communities.

Besides synthetic graph generators, Yang and Leskovec~\cite{yang2015defining} proposed the use of real-world graphs with explicit group annotations (e.g., forums in a social network, categories of products, etc.) to infer what they call ``meta-communities'', and use them to evaluate overlapping community detection algorithms. However, a recent study from Hric, Darst and Fortunato~\cite{hric2014community} reveal a loose correspondence between  communities (the authors refer to them as structural communities) and meta-communities.  This result reveals that  algorithms working for structural communities do not work well for finding meta-communities and vice versa, suggesting significantly different underlying characteristics between the two types of communities, which are yet to be identified.

In this regard and to the best of our knowledge, there does not exist a set of generators to generate graphs with meta-communities for community detection algorithm benchmarking. The closest one is the LDBC-SNB data generator~\cite{Erling:2015:LSN:2723372.2742786} which has been provided by the generation of groups of users in the social network. Even though the generation process does not specifically enforce the generation of groups (meta-communities) for benchmarking community detection algorithms, the study from Prat-Pe\'rez and Dominguez-Sal reveals that these groups are more similar to the real meta-communities than those structural communities generated by the LFR benchmark~\cite{Prat-Perez:2014:CSS:2621934.2621942}.

The differences observed between structural and meta-communities reveal the need of more accurate community definitions tight more specifically to the domain or use case. Current community detection algorithms and graph generators for community detection are stuck to the traditional (and vague) definition of community, assuming that there exists a single algorithm that would fit all the use cases. Thus, future work requires the study of domain-specific community characteristics that can be used to generate graphs with a community structure that accurately resembles that of specific use cases, and thus revealing which are the best algorithms for each particular scenario.

### 2018-09-24 generators_general.tex
Sherif Changes

First of all we focus on papers dealing with generation of general graph
data not aiming at a particular application domain. Currently, there exists a
number of tools which involve a kind of general graph data generator, such as
gtools from projects nauty and Traces~\cite{gtools} or the Stanford
GraphBase~\cite{GraphBase}. We, however, focus on primarily
generating/benchmarking projects targeting Big Data world.

@@

\paragraph{GraphGen} For the purpose of testing the scalability of an indexing
technique called FG-index~\cite{Cheng:2007:FTV:1247480.1247574} on the size of the
database of graphs, their average size and average density, the authors have
also implemented a synthetic generator called GraphGen~\cite{GraphGen}. It is
based on the IBM synthetic data generation code for associations and sequential
patterns\footnote{From 1996, no longer available at
\url{http://www.almaden.ibm.com/cs/projects/iis/hdb/Projects/data
mining/mining.shtml}}. GraphGen creates a collection of labeled, undirected and
connected graphs aiming at performance evaluation of frequent subgraph mining
algorithms and graph query processing algorithms. The result is represented as a
list of graphs, each consisting of a list of nodes and a list of edges.

@@

\paragraph{BTER} BTER (Block Two-Level
Erd\"{o}s-R\'{e}ny)~\cite{kolda2014scalable} is a graph generator based on the
creation of multiple Erd\"{o}s-R\'{e}ny graphs with different connection
probabilities, connected randomly between them. As the main feature, BTER is able
to reproduce input degree distributions and average clustering
coefficient per degree values. The generator starts by grouping the vertices
by degree $d$, and forming groups of size $d+1$ of nodes with degree $d$. Then, these
groups are assigned an internal edge probability in order to match the observed
average clustering coefficient of the nodes of such degree. Based on this
probability, for each node, the excess degree (i.e, the degree that in
expectation will not be realized internally in the group) is computed and used to connect
nodes from different groups at random. The authors describe a highly scalable
MapReduce based implementation that is capable of generating large graphs (with
billions of nodes) in reasonable amounts of time.

\paragraph{Darwini} Darwini~\cite{edunov2016darwini} is an extension of BTER
designed to run on Vertex Centric computing frameworks like
Pregel~\cite{malewicz2010pregel} or Apache Giraph~\cite{ching2015one}, with the
additional feature that it is more accurate when reproducing the clustering
coefficient of the input graph. Instead of just focusing on the average
clustering coefficient for each degree, Darwini is able to model the clustering
coefficient distribution per degree. It achieves this by grouping the vertices
of the graph into buckets by the expected number of closed triangles they need
to close in order to attain the expected clustering coefficient, which is
sampled from the input distributions. Then, the vertices in each bucket are
connected randomly with a probability that would produce, in expectation, the
desired number of triangles for such bucket. Then, as in BTER, the excess degree
is used to connect the different buckets. The authors report that Darwini is
able to generate graphs with billions and even trillions of edges.

>>>>

We start by focusing on approaches that have been designed for dealing with the generation of general graph
data that is not aimed at a particular application domain. Currently, there exists a
number of tools which involve a kind of general graph data generator, such as
gtools from projects nauty and Traces~\cite{gtools} or the Stanford
GraphBase~\cite{GraphBase}. We, however, focus on primarily
generating/benchmarking projects targeting the Big Data world.

@@

\paragraph{GraphGen} For the purpose of testing the scalability of an indexing
technique called FG-index~\cite{Cheng:2007:FTV:1247480.1247574} on the size of the
database of graphs, their average size and average density, the authors have
also implemented a synthetic generator called GraphGen~\cite{GraphGen}. It is
based on the IBM synthetic data generation code for associations and sequential
patterns\footnote{From 1996, no longer available at
\url{http://www.almaden.ibm.com/cs/projects/iis/hdb/Projects/data
mining/mining.shtml}}. GraphGen creates a collection of labeled, undirected and
connected graphs which focuses on the performance evaluation of frequent subgraph mining
algorithms and graph query processing algorithms. The result is represented as a
list of graphs, each consisting of a list of nodes and a list of edges.

@@

\paragraph{BTER} BTER (Block Two-Level
Erd\"{o}s-R\'{e}ny)~\cite{kolda2014scalable} is a graph generator based on the
creation of multiple Erd\"{o}s-R\'{e}ny graphs with different connection
probabilities  of which they are connected randomly between them. As the main feature, BTER is able
to reproduce input degree distributions and average clustering
coefficient per degree values. The generator starts by grouping the vertices
by degree $d$, and forming groups of size $d+1$ of nodes with degree $d$. Then, these
groups are assigned an internal edge probability in order to match the observed
average clustering coefficient of the nodes of such degree. Based on this
probability, for each node, the excess degree (i.e, the degree that in
expectation will not be realized internally in the group) is computed and used to connect
nodes from different groups at random. The authors describe a highly scalable
MapReduce based implementation that is capable of generating large graphs (with
billions of nodes) in a reasonable amounts of time.

\paragraph{Darwini} Darwini~\cite{edunov2016darwini} is an extension of BTER
designed to run on Vertex Centric computing frameworks like
Pregel~\cite{malewicz2010pregel} or Apache Giraph~\cite{ching2015one}, with the
additional feature that it is more accurate when reproducing the clustering
coefficient of the input graph. Instead of just focusing on the average
clustering coefficient for each degree, Darwini is able to model the clustering
coefficient distribution per degree. It achieves this by grouping the vertices
of the graph into buckets by the expected number of closed triangles that they need
to close in order to attain the expected clustering coefficient, which is
sampled from the input distributions. Then, the vertices in each bucket are
connected randomly with a probability that would produce the expected 
desired number of triangles for such bucket. Then, as in BTER, the excess degree
is used to connect the different buckets. The authors report that Darwini is
able to generate graphs with billions and even trillions of edges.

### 2018-09-24 generators_streaming.tex
Sherif Changes

\subsubsection{Graph Data Streaming}
\label{sec:generators_streaming}
One way for dealing with big graphs is to process them using the \emph{streaming} mode where the data stream could consist of the edges of the graph. In this mode, the graph processing algorithms can process the input stream in the order it arrives while using only a limited amount memory~\cite{mcgregor2014graph}. The streaming mode has mainly  attracted the attention of the RDF and Semantic Web community. Thus, Phuoc et al.~\cite{le2012linked} presented an evaluation framework for linked stream data processing engines. The framework uses a data generator for Stream Social network data Generator (S2Gen) that simulates what users continuously generate on their social network activities (e.g., posts) in addition to the  user metadata such as users' profile information, social network relationships, posts, photos and GPS information. The data generator of this framework provides the users the flexibility to control the characteristics of the generated data by tanning a range of parameters including the period in which the social activities are generated (generating period), the maximum number of posts/comments/photos for each user per week and the correlation probabilities between the different objects (e.g., users) in the social network.

Tommasini et al.~\cite{tommasini2017rsplab} introduced another framework for benchmarking RDF Stream Processing systems, RSPLab. The Streamer component of this framework is designed to publish RDF streams from the various existing RDF benchmarks (e.g., BSBM, LUBM) (see Section~\ref{sec:generators_LinkedData}). In particular, the Streamer  component uses TripleWave\footnote{\url{http://streamreasoning.github.io/TripleWave/}}, an open-source framework for publishing and sharing RDF streams on the Web~\cite{mauri2016triplewave}.   TripleWave acts as a mean for plugging in diverse Web data sources and for consuming streams in both push and pull mode.

>>>>

\subsubsection{Graph Data Streaming}
\label{sec:generators_streaming}
One way for dealing with big graphs is to process them using the \emph{streaming} mode where the data stream could consist of the edges of the graph. In this mode, the graph processing algorithms can process the input stream in the order it arrives while using only a limited amount of memory~\cite{mcgregor2014graph}. The streaming mode has mainly  attracted the attention of the RDF and Semantic Web community. Thus, Phuoc et al.~\cite{le2012linked} presented an evaluation framework for linked stream data processing engines. The framework uses a data generator for the Stream Social network data Generator (S2Gen) that simulates what users continuously generate on their social network activities (e.g., posts) in addition to the  user metadata such as users' profile information, social network relationships, posts, photos and GPS information. The data generator of this framework provides the users the flexibility to control the characteristics of the generated data by tanning a range of parameters including the period in which the social activities are generated (generating period), the maximum number of posts/comments/photos for each user per week and the correlation probabilities between the different objects (e.g., users) in the social network.

Tommasini et al.~\cite{tommasini2017rsplab} introduced another framework for benchmarking RDF Stream Processing systems, RSPLab. The Streamer component of this framework is designed to publish RDF streams from the various existing RDF benchmarks (e.g., BSBM, LUBM) (see Section~\ref{sec:generators_LinkedData}). In particular, the Streamer  component uses TripleWave\footnote{\url{http://streamreasoning.github.io/TripleWave/}}, an open-source framework for publishing and sharing RDF streams on the Web~\cite{mauri2016triplewave}.   TripleWave acts as a mean for plugging-in diverse Web data sources and for consuming streams in both push and pull mode.

### 2018-09-24 introduction.tex
Sherif Changes

To the best of our knowledge, our work is the first surveying the broad
landscape of graph data generators spanning different data-intensive applications and
targeting many computer science subfields. In particular, we cover graph database
generators, graph processing generators, social networks and community detection
generators, Semantic Web data generators, graph
analytics generators, IoT, Telecommunication and graph streaming generators,
and Machine Learning and graph mining generators. A comprehensive study of
graph generators is yet missing for many of the specific subfields mentioned
above.

A limited subset of graph database
generators, parallel and distributed graph processing generators, along
with a few of the Semantic Web data generators presented in our survey have
been discussed in a related book chapter
\cite{BFHI18} while cross-comparing them with respect to input, output,
supported workload, data model and query language, along with the
distinguished chokepoints. However, the provided classification is
inherently database-oriented.
In our work, we provide a more comprehensive and broader classification
 serving the purpose of letting any
researcher or practitioner interested in data generation make a wise
choice of the desired graph
generator based on its functional and goal-driven features (such as the application domain,
the supported operations and the key configuration options).
Moreover, in contrast with \cite{BFHI18}, our work encompasses graph generators of several
diverse communities, not limiting its scope to a few generators of the database
and graph processing communities.

@@

Aggarwal and Subbian \cite{AggarwalS14} have surveyed evolution analysis in
graphs, by primarily focusing on data mining maintenance methods and on analytical
quantification and explanation of the changes of the underlying networks.
A brief discussion on evolutionary
network data generators is carried out in
the paper. The data generation of evolutionary networks based on
Densification Power Law (DPL) and shrinking diameters \cite{LeskovecKF05} and community-guided
attachment properties \cite{LeskovecKF05} is then considered, along with tackling Kronecker
recursion with recursive tensor multiplication \cite{AkogluMF08}.
We refer the reader to this survey for evolutionary network generators and
we further discuss the open challenges of evolving graph data in Section 5.
%\TODO{Better highlight the differences when Sherif's part is in.}

@@

\paragraph*{Outline} The rest of the text is structured as follows:
In Section~\ref{sec:comparison} we provide the opening categorization and comparison of the generators. In Section~\ref{sec:generators} we overview the existing graph data generators and their main features in the frame of the proposed categories. To provide a complete overview, in Section~\ref{sec:data_sets} we list most commonly used testing data sets for the particular categories.  In Section~\ref{sec:challenges} we discuss challenges and open problems of graph data synthesis and in Section~\ref{sec:conclusion} we conclude.

>>>>

To the best of our knowledge, our work is the first to survey the broad
landscape of graph data generators spanning different data-intensive applications and
targeting many computer science subfields. In particular, we cover graph database
generators, graph processing generators, social networks and community detection
generators, Semantic Web data generators, graph
analytics generators, IoT, Telecommunication and graph streaming generators,
and Machine Learning and graph mining generators. However, the literature is still lacking a comprehensive study of
graph generators  for many of the specific subfields mentioned above.

A limited subset of graph database
generators, parallel and distributed graph processing generators, along
with a few of the Semantic Web data generators presented in our survey have
been discussed in a related book chapter
\cite{BFHI18} while cross-comparing them with respect to input, output,
supported workload, data model and query language, along with the
distinguished chokepoints. However, the provided classification is
inherently database-oriented.
In our work, we provide a more comprehensive and broader classification
 that serves the purpose of letting any
researcher or practitioner interested in data generation  to be able to make a better 
choice of the desired graph
generator based on its functional and goal-driven features (such as the application domain,
the supported operations and the key configuration options).
Moreover, in contrast with \cite{BFHI18}, our work encompasses graph generators of several
diverse communities, not limiting its scope to a few generators of the database
and graph processing communities.

@@

Aggarwal and Subbian \cite{AggarwalS14} have surveyed the evolution of analysis in
graphs, by primarily focusing on data mining maintenance methods and on analytical
quantification and explanation of the changes of the underlying networks.
A brief discussion on evolutionary
network data generators is carried out in
the paper. The data generation of evolutionary networks is based on
Densification Power Law (DPL) and shrinking diameters \cite{LeskovecKF05} and community-guided
attachment properties \cite{LeskovecKF05} is then considered, along with tackling Kronecker
recursion with recursive tensor multiplication \cite{AkogluMF08}.
We refer the reader to this survey for evolutionary network generators and
we further discuss the open challenges of evolving graph data in Section 5.
%\TODO{Better highlight the differences when Sherif's part is in.}

@@

\paragraph*{Outline} The rest of this paper is organized as follows: Section~\ref{sec:comparison}  provides the opening categorization and comparison of the generators. Section~\ref{sec:generators} provides an overview of the existing graph data generators and their main features in the frame of the proposed categories. To provide a complete overview,  we list the most commonly used testing data sets for the particular categories in Section~\ref{sec:data_sets}.  In Section~\ref{sec:challenges}, we highlight some of the challenges and open problems of graph data synthesis before we conclude in Section~\ref{sec:conclusion}.
